---
title: "DataFest 2020: COVID-19 Analysis"
author: "Team Half-Git Up: Alyssa Shi and Drew Flanagan"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE, 
                      message=FALSE)
```

```{r load-packages}
library(tidyverse)
library(broom)
library(knitr) 
library(dplyr)
library(lubridate)
library(extrafont)
```

```{r load-data}
mobility <- read_csv("Data/Global_Mobility_Report.csv")
covid19 <- read_csv("Data/us_states_covid19_daily.csv")
covidus <- read_csv("Data/us_covid19_daily.csv")
states <- read_csv("Data/StatesFIPSCodes.csv")
```

Data Sources:
https://www.google.com/covid19/mobility/


https://www.kaggle.com/sudalairajkumar/covid19-in-usa#us_states_covid19_daily.csv


https://data.world/uscensusbureau/fips-state-codes



### Data Wrangling

Muating states to be able to join other datasets:

```{r}
states <- states %>%
  select(STUSAB, STATE_NAME) %>%
  mutate(state = STUSAB,
         sub_region_1 = STATE_NAME) %>%
  select(state, sub_region_1)
```

Joining: 

```{r}
covid19 <- left_join(covid19, states)
```

Joining USA data with state data:

```{r}
covidus <- covidus %>%
  mutate(state = "USA") %>%
  select(date, state, positive)

COVID <- full_join(covid19, covidus) %>%
  mutate(date = ymd(date),
         sub_region_1 = if_else(is.na(sub_region_1), "USA", sub_region_1)) %>%
  select(date, state, positive, negative, sub_region_1)
```

Altering mobility, imputing 0 for missing values:

```{r}
mobility <- mobility %>%
  filter(country_region_code == "US") %>%
  mutate(date = ymd(date),
         sub_region_1 = if_else(is.na(sub_region_1), "USA", sub_region_1),
         retail_rec = if_else(is.na(retail_and_recreation_percent_change_from_baseline), 0, retail_and_recreation_percent_change_from_baseline),
         groc_pharm = if_else(is.na(grocery_and_pharmacy_percent_change_from_baseline), 0, grocery_and_pharmacy_percent_change_from_baseline),
         parks = if_else(is.na(parks_percent_change_from_baseline), 0, parks_percent_change_from_baseline),
         transit = if_else(is.na(transit_stations_percent_change_from_baseline), 0, transit_stations_percent_change_from_baseline),
         workplace = if_else(is.na(workplaces_percent_change_from_baseline), 0, workplaces_percent_change_from_baseline),
         residential = if_else(is.na(residential_percent_change_from_baseline), 0, residential_percent_change_from_baseline)) %>%
  select(sub_region_1, date, retail_rec, groc_pharm, parks, transit, workplace, residential)
```

Aggregating mobility by region and date to find average mobility:

```{r}
mobility2 <- aggregate(mobility, by = list(mobility$sub_region_1, mobility$date), FUN = mean)

mobility2 <- mobility2 %>%
  mutate(date = Group.2, 
         sub_region_1 = Group.1) %>%
  select(sub_region_1, date, retail_rec, groc_pharm, parks, transit, workplace, residential)
```

Joining datasets:

```{r}
covid19_mob <- inner_join(COVID, mobility2)
```

```{r}
glimpse(covid19_mob)
```

Not all of the variables were used in our analysis!

### Visualizations

Filtering for top 3/bottom 3 states:

```{r}
covid19_mob <- covid19_mob %>%
  filter(state == "NY" | state == "NJ" | state == "MA" | state == "AZ" | state == "AL"| state == "RI") %>%
  mutate(sub_region_1 = factor(sub_region_1, levels = c("New York", "New Jersey", "Massachusetts", "Arizona", "Alabama", "Rhode Island")))
```

Creating graph comparing states:

```{r}
covid19_mob %>%
  ggplot(mapping = aes(x = date)) +
  geom_line(aes(y = workplace, color = "Workplace")) +
  geom_line(aes(y = parks, color = "Parks")) +
  geom_line(aes(y = transit, color = "Transit")) +
  geom_line(aes(y = retail_rec, color = "Retail and Recreation")) +
  geom_line(aes(y = groc_pharm, color = "Grocery and Pharmacy")) +
  geom_line(aes(y = residential, color = "Residential")) +
  facet_wrap(vars(sub_region_1)) +
  labs(title = "Change in Mobility Over Time in States Most and Least Affected by COVID-19", 
       x = "Date", 
       y = "Percent Change in Mobility", 
       color = "Type of Location") +
  theme_bw()
```

Creating graph for New York:

```{r fig.height = 4, fig.width = 8}
covid19_mob %>%
  filter(state == "NY") %>%
  ggplot(mapping = aes(x = date)) +
  geom_line(aes(y = workplace, color = "Workplace")) +
  geom_line(aes(y = parks, color = "Parks")) +
  geom_line(aes(y = transit, color = "Transit")) +
  geom_line(aes(y = retail_rec, color = "Retail and Recreation")) +
  geom_line(aes(y = groc_pharm, color = "Grocery and Pharmacy")) +
  geom_line(aes(y = residential, color = "Residential")) +
  labs(title = "Change in Mobility Over Time in New York", 
       x = "Date", 
       y = "Percent Change in Mobility",
       color = "Type of Location") +
  geom_vline(xintercept = as.numeric(ymd("2020-03-22")), linetype = 4) +
  theme_bw()
```
